Paper

### 2020-01-14 Paper_resubmit.Rmd
last bits and pieces for paper

featureStability = rowSums(coefs[[select.alpha]][,-1])/(splits*folds)
hazards = apply( exp(t(dysplasia.df[, rownames(coefs[[select.alpha]])]) *  coefs[[select.alpha]][,1]), 1, function(x) {
  sd(x)/mean(x)
})

@@

```{r relrisk, echo=F, fig.width=6, fig.height=6}
unadjRR = ggplot(pg.samp, aes(RR)) + geom_histogram(aes(fill=..x..), bins=10, show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(y='n Samples', x='Relative Risk', title='Unadjusted relative risk') + plot.theme
unadjRR
ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR, width=6, height=5, units='in', dpi=300)

@@

## count arms differently than segs??
df2 = df %>% dplyr::group_by(Status, Patient, Endoscopy.Year) %>% dplyr::summarise( 
  'max.Pred'=max(Prediction),
  'max.RR'=max(RR),
  'max.Path'=max(Pathology),
  'CNA'=length(which(seg.type != 'arms' & cn %in% c('loss','gain'))),
  'arms'=length(which(seg.type == 'arms' & cn %in% c('loss','gain'))),
  'ratio'=length(which(seg.type != 'arms' & cn %in% c('loss','gain')))/length(cn)
  ) %>% mutate(Patient = factor(Patient))
#as.data.frame(df2)
#df2 = subset(df2, max.Path %in% c('BE','ID','LGD'))

@@

el = df2 %>% filter(early | late)
el2 = el %>% dplyr::group_by(Status, Patient) %>% dplyr::summarise( diff(arms), diff(CNA) )
wtA = round(wilcox.test(subset(el2, Status == 'Progressor')$`diff(arms)`, subset(el2, Status == 'Non-Progressor')$`diff(arms)`)$p.value, 3)
if (wtA < 0.0001) wtA = '< 0.0001'

p = ggplot(df2 %>% filter(early|late), aes(Endoscopy.Year, arms, color=Patient)) + facet_grid(~Status) +
  geom_point() +  geom_line(data=subset(df2, early | late)) + theme(legend.position = 'none', text=element_text(size=11)) + 
  plot.theme + labs(x='Endoscopy Year', y='CN arm values', title='Initial vs final endoscopy', subtitle=paste('p', wtA))
p
ggsave(filename='plots/status_hetero.png', plot=p, width=8, height=6, units='in', dpi=300)

wtA = round(wilcox.test(subset(df2, early &  Status == 'Progressor')$arms, subset(df2, early & Status == 'Non-Progressor')$arms)$p.value, 3)
wtCN = round(wilcox.test(subset(df2, early & Status == 'Progressor')$CNA, subset(df2, early & Status == 'Non-Progressor')$CNA)$p.value, 2)

m = melt(subset(df2, early), measure.vars=c('CNA','arms'), id.vars=c('Status'))
p = ggplot(m, aes(Status, value, fill=Status)) + facet_wrap(~variable, scales='free_y') + 
  geom_boxplot(outlier.colour = NA) + geom_jitter(width=0.2) + 
  scale_fill_brewer(palette = 'Dark2') + 
  plot.theme + theme(legend.position = 'none', text=element_text(size=11)) +
  labs(y='', x='', title='CN values, initial endoscopy', subtitle=paste('Arms p=',wtA, '   CNA p=',wtCN, sep=''))
p
ggsave(filename='plots/hetero_initial.png', plot=p, width=6, height=4, units='in', dpi=300)

@@

rp = plot.risk.by.path(pg.samp %>% ungroup)
rp
ggsave(filename='plots/pred_by_path.png', plot=(rp + theme(legend.position = 'none')), width=7, height=6, units='in', dpi=300)
```

@@

```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=6}
all.preds = pg.samp %>% dplyr::group_by(Status, Patient, Risk) %>% tally %>% spread(Risk, n) %>% ungroup %>%
  mutate_if(is.numeric, funs( ifelse(is.na(.), 0, .) )) %>% 
  left_join(pg.samp %>% dplyr::group_by(Status, Patient) %>% tally, by=c('Status','Patient')) %>%
  dplyr::rename('n.samples'='n') %>% 
  dplyr::mutate(Patient = factor(Patient))

@@

rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) +
  geom_bar(stat='identity', fill=BarrettsProgressionRisk:::riskColors()[['High']], show.legend = F) + 
  geom_errorbar(color='grey55', width=0.2) +
  geom_text(aes(label=paste(round(pred.r*100), '%', sep='')), color='black', nudge_y=2) +
  geom_text(aes(y=5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white', nudge_y=2) + 
  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') + coord_cartesian(ylim = c(0,100)) + 
  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12) ) 
rp
ggsave(filename='plots/pred_endo.png', plot=rp, width=6, height=5, units='in', dpi=300)

@@

@@

```{r}
rp = plot.risk.by.path(val.preds) + labs(title='Validation samples classified by pathology')  
rp
ggsave(filename='plots/validation/val_pred_by_path.png', plot=rp, width=5, height=6, units='in', dpi=300)

@@

sfit = survfit( Surv(months.before.final, Diagnosis, type='right') ~ Risk, data=merged, conf.int=0.95, conf.type='log', na.action=na.omit)
names(sfit$strata) = sub('.*=','', names(sfit$strata))

kmp = survminer::ggsurvplot(sfit, size=1, linetype="strata",risk.table=T, risk.table.col='strata',
                      conf.int=T, censor=T, legend='none', 
                      title='Risks over time', 
                      subtitle='Training & Validation predictions',
                      xlab='Time (months prior to final sample)',
                      ylab = 'Progression risk',
                      ggtheme=theme_minimal(), palette=unlist(BarrettsProgressionRisk::riskColors()), font.legend=c(14,'bold'))

@@

p = ggplot( x %>% filter(!is.na(cv.performance_mean)), aes(x=n.patients, y=cv.performance_mean, group=n.patients, fill=Cohort)) + 
  facet_grid(~Cohort, scales = 'free_x', space='free_x') + 
  geom_col() + geom_errorbar(aes(ymin=cv.performance_mean-cv.performance_sd/2, ymax=cv.performance_mean+cv.performance_sd/2), width=0.2, color='grey39') +
  geom_text(aes(label=round(cv.performance_mean,2)), y=0.05) + 
  scale_y_continuous(expand=c(0,0), limits=c(0,0.8)) + 
  scale_fill_brewer(palette = 'Set2',name='') + 
  plot.theme + theme(legend.position = 'bottom', strip.text = element_blank(), panel.spacing = unit(0.5, 'lines')) +
  labs(title='Cross validation performance', y='', x='# patients used to train')

@@

p1 = ggplot(x %>% filter(!is.na(n.coefs_mean)), aes(x=n.patients, y=n.coefs_mean, group=n.patients, fill=Cohort)) + 
  facet_grid(~Cohort, scales = 'free_x', space='free_x') + 
  geom_col() + geom_errorbar(aes(ymin=n.coefs_mean-n.coefs_sd/2, ymax=n.coefs_mean+n.coefs_sd/2), width=0.2, color='grey39') +
  geom_text(aes(label=round(n.coefs_mean,2)), y=5) +
  scale_y_continuous(expand=c(0,0)) + scale_fill_brewer(palette = 'Set2') + 
  plot.theme + theme(legend.position = 'none', strip.text = element_blank(), panel.spacing = unit(0.5, 'lines')) +
  labs(title='# Coefficients',y='', x='# patients used to train')

p2 = ggplot(x, aes(x=n.patients, y=auc_mean, group=n.patients, fill=Cohort)) +
  facet_grid(~Cohort, scales = 'free_x', space='free_x') + 
  geom_col() + 
  geom_errorbar(aes(ymin=auc_mean-auc_sd/2, ymax=auc_mean+auc_sd/2), width=0.2, color='grey39') +
  geom_text(aes(label=round(auc_mean,2)), y=0.05) +
  geom_text( data=filter(x,Cohort == 'sWGS (all) + SNP'), aes(label=Type, y=0.1) ) +
  scale_y_continuous(expand=c(0,0), limits=c(0,1)) + scale_fill_brewer(palette = 'Set2') + 
  plot.theme + theme(legend.position = 'none', strip.text = element_blank(), panel.spacing = unit(0.5, 'lines')) +
  labs(title='AUC', y='', x='# patients used to train')

>>>>

featureStability = rowSums(coefs[[select.alpha]][,-1])/(splits*folds)
hazards = apply( exp( t(dysplasia.df[,rownames(coefs[[select.alpha]])]) * coefs[[select.alpha]][,1]), 1, function(x) {
  sd(x)/mean(x)
})

@@

```{r relrisk, echo=F, fig.width=6, fig.height=6}
unadjRR = ggplot(pg.samp, aes(RR)) + geom_histogram(aes(fill=..x..), bins=10, show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(y='n Samples', x='log(Relative Risk)', title='Unadjusted relative risk') + plot.theme
unadjRR
ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR, width=6, height=5, units='in', dpi=300)

@@

## count arms differently than segs??
df2 = df %>% dplyr::group_by(Status, Patient, Endoscopy.Year) %>% dplyr::summarise( 
  'max.Pred'=max(Prediction),
  'max.RR'=max(RR),
  'max.Path'=max(Pathology),
  'CNA'=length(which(seg.type != 'arms' & cn %in% c('loss','gain'))),
  'arms'=length(which(seg.type == 'arms' & cn %in% c('loss','gain'))),
  'ratio'=length(which(seg.type != 'arms' & cn %in% c('loss','gain')))/length(cn)
  ) %>% ungroup %>% mutate(Patient = factor(Patient))
#as.data.frame(df2)
#df2 = subset(df2, max.Path %in% c('BE','ID','LGD'))

@@

el = df2 %>% filter(early | late)
el2 = el %>% dplyr::group_by(Status, Patient) %>% dplyr::summarise( diff(arms), diff(CNA) )
wtA = signif(wilcox.test(subset(el2, Status == 'Progressor')$`diff(arms)`, subset(el2, Status == 'Non-Progressor')$`diff(arms)`)$p.value,3)
#if (wtA < 0.0001) wtA = '< 0.0001'

p = ggplot(df2 %>% filter(early|late), aes(Endoscopy.Year, arms, color=Patient)) + facet_grid(~Status) +
  geom_point() +  geom_line(data=subset(df2, early | late)) + theme(legend.position = 'none', text=element_text(size=11)) + 
  plot.theme + labs(x='Endoscopy Year', y='CN arm values', title='Initial vs final endoscopy', subtitle=paste('p-value=', wtA))
p
ggsave(filename='plots/status_hetero.png', plot=p, width=8, height=6, units='in', dpi=300)

wtA = signif(wilcox.test(subset(df2, early &  Status == 'Progressor')$arms, subset(df2, early & Status == 'Non-Progressor')$arms)$p.value, 3)
wtCN = signif(wilcox.test(subset(df2, early & Status == 'Progressor')$CNA, subset(df2, early & Status == 'Non-Progressor')$CNA)$p.value, 2)

m = melt(subset(df2, early), measure.vars=c('CNA','arms'), id.vars=c('Status'))
p = ggplot(m, aes(Status, value, fill=Status)) + facet_wrap(~variable, scales='free_y') + 
  geom_boxplot(outlier.colour = NA) + geom_jitter(width=0.2) + 
  scale_fill_brewer(palette = 'Dark2') + 
  plot.theme + theme(legend.position = 'none', text=element_text(size=11)) +
  labs( y='Segment count', title='CN values, initial endoscopy', subtitle=paste('Arms p=',wtA, '   CNA p=',wtCN, sep=''))
p
ggsave(filename='plots/hetero_initial.png', plot=p, width=6, height=4, units='in', dpi=300)

@@

rp = plot.risk.by.path(pg.samp %>% ungroup) + labs(title='Samples classified by pathology', y='% Classified')

ggsave(filename='plots/pred_by_path.png', plot=(rp + theme(legend.position = 'none')), width=7, height=6, units='in', dpi=300)
```

@@

```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=6, eval=F}
all.preds = pg.samp %>% dplyr::group_by(Status, Patient, Risk) %>% tally %>% spread(Risk, n) %>% ungroup %>%
  mutate_if(is.numeric, funs( ifelse(is.na(.), 0, .) )) %>% 
  left_join(pg.samp %>% dplyr::group_by(Status, Patient) %>% tally, by=c('Status','Patient')) %>%
  dplyr::rename('n.samples'='n') %>% 
  dplyr::mutate(Patient = factor(Patient))

@@

rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) +
  geom_bar(stat='identity', fill=BarrettsProgressionRisk:::riskColors()[['High']], show.legend = F) + 
  geom_errorbar(color='grey55', width=0.2) +
  geom_text(aes(label=paste(round(pred.r*100), '%', sep='')), color='black', nudge_y=2) +
  geom_text(aes(y=5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white', nudge_y=2) + 
  labs(x='Months prior to endpoint', y='Percent classified', title='High risk classifications prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') + coord_cartesian(ylim = c(0,100)) + 
  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12) ) 
rp
ggsave(filename='plots/pred_endo.png', plot=rp, width=6, height=5, units='in', dpi=300)

@@

vdir ='~/Data/BarrettsProgressionRisk/Analysis/validation/pcf_perPatient/50kb'
residuals = do.call(bind_rows, purrr::map(list.files(vdir,'residuals.tsv',recursive = T, full.names = T), function(f) read_tsv(f, col_types = 'ccddddl')))
failedQC = filter(residuals, varMAD_median > 0.008)
```

@@

```{r}
rp = plot.risk.by.path(val.preds) + labs(title='Validation samples classified by pathology', y='% Classified')  
rp
ggsave(filename='plots/validation/val_pred_by_path.png', plot=rp, width=5, height=6, units='in', dpi=300)

@@

#merged = merged %>% mutate(Diagnosis = ifelse(Status == 'P' & months.before.final == 0, 1, 0))

merged %>%

sfit = survfit( Surv(months.before.final, Diagnosis) ~ Risk, data=merged, conf.int=0.95, conf.type='log', na.action=na.omit)
names(sfit$strata) = sub('.*=','', names(sfit$strata))

kmp = 
  survminer::ggsurvplot(sfit, size=1, linetype="strata",risk.table=F, risk.table.col='strata',
                      conf.int=T, censor=T, legend='none',# xlim=c(0,150),
                      title='Risk Classifications', 
                      subtitle='Training & Validation cohorts',
                      xlab='Months prior to final sample',
                      ylab = 'Progression free fraction',
                      ggtheme=theme_minimal(), palette=unlist(BarrettsProgressionRisk::riskColors()), font.legend=c(14,'bold'))

kmp$plot = kmp$plot + scale_x_reverse() 
kmp$table = kmp$table + scale_x_reverse()

@@

p = ggplot( x %>% filter(!is.na(cv.performance_mean)), aes(x=n.patients, y=cv.performance_mean, group=n.patients, fill=Cohort)) + 
  facet_grid(~Cohort, scales = 'free_x', space='free_x') + 
  geom_col() + geom_errorbar(aes(ymin=cv.performance_mean-cv.performance_sd/2, ymax=cv.performance_mean+cv.performance_sd/2), width=0.2, color='grey39') +
  geom_text(aes(label=round(cv.performance_mean,2)), y=0.05) + 
  scale_y_continuous(expand=c(0,0), limits=c(0,0.8)) + 
  scale_fill_brewer(palette = 'Set2',name='') + 
  plot.theme + theme(legend.position = 'bottom', strip.text = element_blank(), panel.spacing = unit(0.5, 'lines')) +
  labs(title='Cross validation performance', y='', x='# patients')

@@

p1 = ggplot(x %>% filter(!is.na(n.coefs_mean)), aes(x=n.patients, y=n.coefs_mean, group=n.patients, fill=Cohort)) + 
  facet_grid(~Cohort, scales = 'free_x', space='free_x') + 
  geom_col() + geom_errorbar(aes(ymin=n.coefs_mean-n.coefs_sd/2, ymax=n.coefs_mean+n.coefs_sd/2), width=0.2, color='grey39') +
  geom_text(aes(label=round(n.coefs_mean,2)), y=5) +
  scale_y_continuous(expand=c(0,0)) + scale_fill_brewer(palette = 'Set2') + 
  plot.theme + theme(legend.position = 'none', strip.text = element_blank(), panel.spacing = unit(0.5, 'lines')) +
  labs(title='# Coefficients',y='', x='# patients')

p2 = ggplot(x, aes(x=n.patients, y=auc_mean, group=n.patients, fill=Cohort)) +
  facet_grid(~Cohort, scales = 'free_x', space='free_x') + 
  geom_col() + 
  geom_errorbar(aes(ymin=auc_mean-auc_sd/2, ymax=auc_mean+auc_sd/2), width=0.2, color='grey39') +
  geom_text(aes(label=round(auc_mean,2)), y=0.05) +
  geom_text( data=filter(x,Cohort == 'sWGS (all) + SNP'), aes(label=Type, y=0.1) ) +
  scale_y_continuous(expand=c(0,0), limits=c(0,1)) + scale_fill_brewer(palette = 'Set2') + 
  plot.theme + theme(legend.position = 'none', strip.text = element_blank(), panel.spacing = unit(0.5, 'lines')) +
  labs(title='AUC', y='', x='# patients')
